% !TEX root = hw1.tex
\documentclass{article}

\input{preamble.tex}

\title{Homework 1}
\author{Mihai Codoban \and Arpit Christi \and Caius Brindescu \and Morgan Shirley}

\begin{document}

\maketitle

\input{problem1}

\problem{2}

\problem{3}

A longest palindrome subsequence of a string is a longest common subsequence of that string and a reverse-order copy of the string. The algorithm for finding the length of a longest common subsequence of two strings is given below.

\begin{algorithm}
	\caption{Length of LCS with Dynamic Programming}
	\begin{algorithmic}

	\Function{LCS}{$x[1..n]$, $y[1..m]$, T$[1..n][1..m]$}

		\For{$i=0 \to n$}
		
			\For{$j=0 \to m$}

				\If {$i=0$} \State T$[i][j] \gets 0$

				\ElsIf {$j=0$} \State T$[i][j] \gets 0$

				\ElsIf {$x[i] = y[j]$} \State T$[i][j] \gets$ T$[i-1][j-1] + 1$

				\Else \State T$[i][j] \gets max($T$[i-1][j], $T$[i, j-1])$

				\EndIf

			\EndFor

		\EndFor

	  \EndFunction
	\end{algorithmic}
\end{algorithm}


The fact that the length of a longest palindromic subsequence is less than or equal to the length of a longest common subsequence of a string and its reverse is obvious, because it is a common subsequence of the two by the definition of palindrome. To show that the length of an LPS of a string is strictly equal to the length of an LCS of the string and its reverse, consider such an LCS. We can construct an LPS by taking the middle of the LCS and replacing the second half of the subsequence with the reverse of the first half of the subsequence. This new palindromic subsequence has the same length as the LCS and is present in the original string; to see this, note that we are given the fact that the first half of the subsequence is present in both strings, which means that it must be present \emph{in reverse} in both strings after a certain pivot point. Therefore, this must be an LPS of the original string.

The algorithm creates a dynamic programming table of the solutions for smaller substrings of $x$ and $y$. Calculating the value for each step takes constant time, and since there are $m$ rows and $n$ columns the algorithm takes $O(n*m)$ time to complete. Creating the reverse of the string will take $O(n)$ time and the reverse will have equal length to the original string, so overall the longest palindromic subsequence algorithm takes $O(n) + O(n^2) = O(n^2)$ time to complete.

\problem{4}

\end{document}